---
output:
  pdf_document: default
  html_document: default
---
# LIBRARIES

```{r}
library(tidyverse)
library(dplyr)
library(DataExplorer)
library(ggplot2)
library(ggthemes)
library(wesanderson)
library(caTools)# spliting dataset
library(caret)# partition, confusion matrix, calculate F1, precision...
library(e1071)# for Naieve Bayes train model
library(ROSE)# for ROC curve
library(rattle)# fancy plot
library(ggpubr)
library(rpart)
library(rpart.plot)
library(randomForest)
library(irr)
library("psych")
library(class)
```


# DATA EXPLORATION

```{r}
library(readr)

census_in <- read_csv("/Users/chiarasaini/Desktop/uni/SL lectures/SL new/Statistical learning project - Chiara Saini /Supervised/adult.csv") #Set path
head(census_in)
```


## Check for NA values
```{r}
colSums(is.na(census_in))
```

```{r}
nrow(census_in)
```


```{r}
str(census_in)
summary(census_in)
```


## Group and count each group

```{r}
# Workclass
census_in %>% group_by(workclass) %>% count()
# Education
census_in %>% group_by(education) %>% count()
# Marital status
census_in %>% group_by(marital.status) %>% count()
# Occupation
census_in %>% group_by(occupation) %>% count()
# Relationship
census_in %>% group_by(relationship) %>% count()
# Race
census_in %>% group_by(race) %>% count()
# Sex
census_in %>% group_by(sex) %>% count()
# Native country
census_in %>% group_by(native.country) %>% count()
# Income
census_in %>% group_by(income) %>% count()

```

# DATA MANIPULATION

```{r}
# Change  ? to NA value
head(census_in)
census_in[census_in == "?"]<- NA

colSums(is.na(census_in))
```

```{r}
# Omit missing data records
census<- na.omit(census_in)
head(census)
colSums(is.na(census))
```

```{r}
#mark income<=50K as 0 and income>50 as 1

#census$income<-ifelse(census$income=='<=50K',0,1)
```

```{r}
# Grouping for workclass attribute
unempl <- function(job_role){
    job_role <- as.character(job_role)
    if (job_role=="Never-worked" | job_role=="Without-pay"){
        return("Other")
    }else{
        return(job_role)
    }
}
census$workclass <- sapply(census$workclass,unempl)
table(census$workclass)
```

```{r}
# Grouping for workclass attribute
grp_emp <- function(job_role){
    if (job_role=="Local-gov" | job_role=="State-gov" | job_role=="Federal-gov"){
        return("Government")
    }else if (job_role=="Self-emp-inc" | job_role=="Self-emp-not-inc"){
        return("Self_employ")
    }else{
        return(job_role)
    }
}
census$workclass <- sapply(census$workclass,grp_emp)
table(census$workclass)
```

```{r}
# Grouping for education attribute
educ <- function(educ_level){
    if (educ_level=="Preschool" | educ_level=="1st-4th" | educ_level=="5th-6th"| educ_level=="7th-8th"| educ_level=="9th"| educ_level=="10th"| educ_level=="11th"| educ_level=="12th"){
        return("not_HS_grad")
    }else{
        return(educ_level) }
    }
census$education <- sapply(census$education,educ)
table(census$education)
```

```{r}
# Grouping for marital.status attribute

group_marital <- function(marital_stat){
    marital_stat <- as.character(marital_stat)
    
    # Not-Married
    if (marital_stat=="Separated" | marital_stat=="Divorced" | marital_stat=="Widowed"){
        return("Not-Married")
    
    # Never-Married   
    }else if(marital_stat=="Never-married"){
        return(marital_stat)
    
     #Married
    }else{
        return("Married")
    }
}

census$marital.status <- sapply(census$marital.status,group_marital)
table(census$marital.status)
```

```{r}
# Grouping for native.country attribute
Asia <- c("China","Hong","India","Iran","Cambodia","Japan", "Laos","Philippines" ,"Vietnam" ,"Taiwan", "Thailand")

N.A <- c("Canada","United-States","Puerto-Rico")

Europe <- c("England","France","Germany" ,"Greece","Holand-Netherlands","Hungary","Ireland","Italy","Poland","Portugal","Scotland"
            ,"Yugoslavia")

S.A <- c("Columbia","Cuba","Dominican-Republic","Ecuador","El-Salvador","Guatemala","Haiti","Honduras","Mexico","Nicaragua"
                   ,"Outlying-US","Peru","Jamaica","Trinadad&Tobago")
Remaining_count <- c("South")
```

```{r}
# Grouping for native.country attribute
grp_cntry <- function(cntry){
    if (cntry %in% Asia){
        return("Asia")
    }else if (cntry %in% N.A){
        return("N.A")
    }else if (cntry %in% Europe){
        return("Europe")
    }else if (cntry %in% S.A){
        return("S.A")
    }else{
        return("Remaining_count")      
    }
}
census$native.country <- sapply(census$native.country,grp_cntry)
table(census$native.country)
```


```{r}
# Grouping for occupation attribute
occup <- function(occup_type){
    if (occup_type=="Craft-repair" | occup_type=="Farming-fishing" | occup_type=="Handlers-cleaners"| occup_type=="Machine-op-inspct"| occup_type=="Transport-moving"){
        return("Worker")
    }
  else if (occup_type=="Other-service" | occup_type=="Priv-house-serv" | occup_type=="Protective-serv" | occup_type=="Tech-support"){
        return("Service")
    }
  else if (occup_type=="Exec-managerial"| occup_type=="Adm-clerical"){
        return("Office")
    }
  else if (occup_type=="Prof-specialty"){
        return("Professional")
    }else{
        return(occup_type) }
    }
census$occupation <- sapply(census$occupation,occup)
table(census$occupation)
```

```{r}
# Converting into factor datatype
census$workclass <- as.factor(census$workclass)
census$native.country <- as.factor(census$native.country)
census$marital.status <- as.factor(census$marital.status)
census$occupation <- as.factor(census$occupation)
census$education <- as.factor(census$education)

str(census)
```




```{r}
#Correlation of numeric independent variables with target varible income
correl <- cor(census%>% select_if(is.numeric))
round(correl, 2)
```
education.num is highly correlated with target(income) and fnwgt is lowest in correlation with target. Therefore during model building it will be dropped.

```{r}
#conventing income to factor
census$income<-as.factor(as.character(census$income))
```

```{r}
# Summary of dataset
summary_census <- psych::describe(census, fast = TRUE)
summary_census

#path_out = '' #Set path
#write.csv(summary_census,paste(path_out,'summary.csv'))
```


# DATA INSIGHT


```{r}
# Income 
table(census$income)
```

```{r}
theme_set(theme_economist())
```

```{r}
df1 <- census %>% 
  group_by(income) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(income))

df1$label <- scales::percent(df1$per)


income_plt <- ggplot(data=df1)+
  geom_bar(aes(x="", y=per, fill=income), stat="identity", width = 1)+
  coord_polar("y", start=0)+
  ggtitle(paste0("Income distribution")) +
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) + 
  scale_fill_manual(values=wes_palette(n=2, name="FantasticFox1"))

income_plt 

#ggsave("income_plt.png", plot = income_plt, path = "") #Set path

```


```{r}

# Age distribution
hist(census$age, 
     main="Age distribution", 
     xlab="Age", 
     ylab="Count",
     border="white", 
     col="#7394D4",
     xlim=c(0,100),
     las=1, 
     breaks=50)

# Comparison age and income
plot1<- ggplot(census, aes(x=age)) + 
  geom_bar(aes(fill=income)) +
  ggtitle(paste0("Comparison between age and income")) + 
  scale_fill_manual(values=wes_palette(n=2, name="FantasticFox1"))


plot1 

#ggsave("plot1.png", plot = plot1, path = "") #Set path
```

```{r}
# Workclass distribution
t1<- table(census$workclass)
barplot(t1,
        main="Workclass distribution",
        xlab = "Workclass", 
        ylab = "Count",
        col = "#7394D4",
        border="white")

# Comparison workclass and income
plot2<- ggplot(census, aes(x=workclass)) + 
  geom_bar(aes(fill=income)) +
  ggtitle(paste0("Comparison between workclass and income")) + 
  scale_fill_manual(values=wes_palette(n=2, name="FantasticFox1"))

plot2 

#ggsave("plot2.png", plot = plot2, path = "") #Set path
```

```{r}
# Comparison between fnlwgt and income
plot12<- ggplot(census, aes(x=fnlwgt)) + 
  geom_density(aes(fill=income), alpha=0.5) +
  ggtitle(paste0("Comparison between final weight and income")) + 
  scale_fill_manual(values=wes_palette(n=2, name="FantasticFox1"))
  
  
plot12 

#ggsave("plot12.png", plot = plot12, path = "") #Set path
```


```{r}
#Education distribution
t2<- table(census$education)
barplot(t2,
        main="Education distribution",
        horiz = T, 
        xlab = "Count",
        ylab = "education",
        col = "#7394D4",
        border="white",
        xlim=c(0,12000),
        cex.names = 0.50,
        las=1)

# Comparison education and income
plot3<- ggplot(census, aes(x=education)) + 
  geom_bar(aes(fill=income)) +
  ggtitle(paste0("Comparison between education and income")) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  scale_fill_manual(values=wes_palette(n=2, name="FantasticFox1"))

plot3

#ggsave("plot3.png", plot = plot3, path = "") #Set path
```


```{r}
# Education number count
t3<- table(census$education.num)
barplot(t3,
        main="Count of education year",
        horiz = T, 
        xlab = "Count", 
        ylab = "Education number (Year)",
        col = "#7394D4",
        border="white",
        las=1)


# Education number and Income
plot4<- ggplot(census, aes(x=education.num)) + 
  geom_bar(aes(fill=income)) +
  ggtitle(paste0("Comparison between years of education and income")) + 
  scale_fill_manual(values=wes_palette(n=2, name="FantasticFox1"))

plot4

#ggsave("plot4.png", plot = plot4, path = "") #Set path
```


```{r}
# Distribution of marital status
t4<- table(census$marital.status)
barplot(t4,
        main="Plot for marital status", 
        xlab = "Marital status", 
        ylab = "Count",
        col = "#7394D4",
        border="white")

# Comparison between marital status and Income 
plot5<- ggplot(census, aes(x=marital.status)) + 
  geom_bar(aes(fill=income)) +
  ggtitle(paste0("Comparison between marital status and income")) + 
  scale_fill_manual(values=wes_palette(n=2, name="FantasticFox1"))

plot5 

#ggsave("plot5.png", plot = plot5, path = "") #Set path
```


```{r}
# Distribution of occupation
t5<- table(census$occupation)
barplot(t5,
        main="Occupation distribution", 
        horiz = T, 
        xlab = "Occupation",
        col = "#7394D4",
        border="white",
        xlim = c(0,12000),
        cex.names = 0.5, 
        las=1)


# Comparison between occupation and income 
plot6<- ggplot(census, aes(x=occupation)) + 
  geom_bar(aes(fill=income)) +
  ggtitle(paste0("Comparison between occupation and income")) + 
  scale_fill_manual(values=wes_palette(n=2, name="FantasticFox1"))

plot6 

#ggsave("plot6.png", plot = plot6, path = "") #Set path
```

```{r}
# Comparison between relationship and income 
plot11<- ggplot(census, aes(x=relationship)) + 
  geom_bar(aes(fill=income)) +
  ggtitle(paste0("Comparison between relationship and income")) + 
  scale_fill_manual(values=wes_palette(n=2, name="FantasticFox1"))+ 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

plot11


#ggsave("plot11.png", plot = plot11, path = "") #Set path
```


```{r}
# Race distribution
t6<- table(census$race)
barplot(t6,
        main = "Race distribution", 
        xlab = "Race",
        ylab ="Count",
        col = "#7394D4",
        border="white",
        ylim=c(0,30000),
        cex.names = 0.6)

# Comparison between race and income
plot7<- ggplot(census, aes(x=race)) + 
  geom_bar(aes(fill=income)) +
  ggtitle(paste0("Comparison between Race and Income")) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ 
  scale_fill_manual(values=wes_palette(n=2, name="FantasticFox1"))
  
plot7

#ggsave("plot7.png", plot = plot7, path = "")#Set path
```


```{r}
# Compare sex and income
t7<- table(census$sex)
barplot(t7,
        main="Sex count", 
        xlab = "Sex",
        ylab = "Count",
        col = "#7394D4",
        border="white")


# Comparison between sex and income
plot8<- ggplot(census, aes(x=sex)) + 
  geom_bar(aes(fill=income)) +
  ggtitle(paste0("Comparison between sex and income")) +
  scale_fill_manual(values=wes_palette(n=2, name="FantasticFox1"))
  
  
plot8

#ggsave("plot8.png", plot = plot8, path = "") #Set path
```


```{r}
# Hours-per-week distribution

plot9<- ggplot(census, aes(x=hours.per.week))+ 
  geom_density(aes(fill=income), alpha = 0.5)  +
  ggtitle(paste0("Comparison between Hours-per-week and Income")) + 
  scale_fill_manual(values=wes_palette(n=2, name="FantasticFox1"))


plot9

#ggsave("plot9.png", plot = plot9, path = "") #Set path
```

```{r}
# Comparison between capital gain and income
plot13<- ggplot(census, aes(x=capital.gain)) + 
  geom_histogram(aes(fill=income)) +
  ggtitle(paste0("Comparison between capital gain and income")) 
  geom_histogram(aes(fill=income), alpha=0.6)
  
  
plot13 <- plot13  + scale_fill_manual(values=wes_palette(n=2, name="FantasticFox1"))

plot13

#ggsave("plot13.png", plot = plot13, path = "") #Set path
```


```{r}
# Comparison between capital loss and income
plot14<- ggplot(census, aes(x=capital.loss)) + 
  geom_histogram(aes(fill=income)) +
  ggtitle(paste0("Comparison between capital loss and income")) 
  geom_histogram(aes(fill=income), alpha=0.6)
  
plot14 <-plot14 + scale_fill_manual(values=wes_palette(n=2, name="FantasticFox1"))
plot14

#ggsave("plot14.png", plot = plot14, path = "") #Set path
```



```{r}

# Comparison between native country and income
plot10<- ggplot(census, aes(x=native.country)) + 
  geom_bar(aes(fill=income)) +
  ggtitle(paste0("Comparison between native country and income")) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  scale_fill_manual(values=wes_palette(n=2, name="FantasticFox1"))
  
  
plot10

#ggsave("plot10.png", plot = plot10, path = "") #Set path
```


# TRAIN, TEST SPLIT
 fnlwgt is not much corrrelated with income, therefore it can be dropped
 
 
 
```{r}
#cato <- c('workclass','education','marital.status','occupation','relationship','race','sex','native.country')
#for(each in cato){
#  census[,each]<-as.integer(census[,each])
#}
```
 

```{r}
#census=census[,-3]
```


```{r}
# Splitting the data 
set.seed(86)
index<-sample(nrow(census),0.70*nrow(census),replace = F)
train<-census[index,]
test<-census[-index,]
```

```{r}
summary(train)
summary(test)
```

```{r}
table(train$income)
table(test$income)
```

```{r}
mydata <- income ~.
```

# RESAMPlING


```{r}
data_balanced_over <- ovun.sample(income~., data=train, method = "both")$data # Resampling to balance data, using method "both"
table(data_balanced_over$income)
train <- data_balanced_over
```
# KNN

```{r}
# create dummy variables for train
dmy.training = dummyVars(" ~ .", data = train)
training.dmy = data.frame(predict(dmy.training, newdata = train))

training.dmy = training.dmy[-48]  # Remove column 48 which is income <=50k
names(training.dmy)[names(training.dmy) == "income..50K"] = "income.more.50k" # Rename the income>50k column


# create dummy variables for test
dmy.test = dummyVars(" ~ .", data = test)
test.dmy = data.frame(predict(dmy.test, newdata = test))

test.dmy = test.dmy[-48]  # Remove column 48 which is income<=50
names(test.dmy)[names(test.dmy) == "income..50K"] = "income.more.50k" # Rename the income>50k column
```

```{r}
head(training.dmy)
head(test.dmy)
```

```{r}
#knn models
knn.training = training.dmy[-47] # Remove target variable
knn.test = test.dmy[-47]

train_labels = census$income[index] 
test_labels = census$income[-index]

# Building the models with different k values
estknn.20 = knn(knn.training, knn.test , train_labels, k=20)
estknn.10 = knn(knn.training, knn.test , train_labels, k=10)
estknn.5 = knn(knn.training, knn.test , train_labels, k=5)

# Confusion Matrix 
cm.knn20 = confusionMatrix(reference = test$income, data = estknn.20, positive = '>50K')
acu.knn20 = confusionMatrix(reference = test$income, data = estknn.20, positive = '>50K')$overall[1]
cm.knn10 = confusionMatrix(reference = test$income, data = estknn.10, positive = '>50K')
acu.knn10 = confusionMatrix(reference = test$income, data = estknn.10, positive = '>50K')$overall[1]
cm.knn5 = confusionMatrix(reference = test$income, data = estknn.5, positive = '>50K')
acu.knn5 = confusionMatrix(reference = test$income, data = estknn.5, positive = '>50K')$overall[1]
```

```{r}
cm.knn20
```

```{r}
cm.knn10
```

```{r}
cm.knn5
```

```{r}
Accuracy<-data.frame(Model=c('K = 20','K = 10','K = 5'),Accuracy=c(acu.knn20,acu.knn10,acu.knn5))


acc_nn <- ggplot(Accuracy,aes(x=Model,y=Accuracy)) +
  geom_bar(stat = 'identity', aes(fill=Model)) +
  ggtitle('Accuracies of Knn')
acc_nn <- acc_nn + scale_fill_manual(values=wes_palette(n=4, name="FantasticFox1"))
acc_nn
#ggsave("acc_nn.png", plot = acc_nn, path = "") #Set path
```



# LOGISTIC REG

```{r}
#mark income<=50K as 0 and income>50 as 1

#census$income<-ifelse(census$income=='<=50K',0,1)
```


```{r}
logic_model <- glm(income ~ ., family = binomial(logit), data = train)
summary(logic_model)
```

```{r}
#calculate p-value of overall Chi-Square statistic
1-pchisq(29268-16768, 21112-21074)
```


```{r}
# Confusion Matrix and predictions

predicted.income = predict(logic_model, newdata=test, type="response")
conf_matrix <- table(test$income, predicted.income > 0.5)
conf_matrix

```
TP = 1828  ->  You predicted positive and it’s true.
TN = 5401  ->  You predicted negative and it’s true.

FP = 389  ->  You predicted positive and it’s false.
FN = 1431   ->  You predicted negative and it’s false.



```{r}
print("Accuracy")
lgAcu <- (1828+5401)/(5401+1828+1431+389)
lgAcu
# Specificity 
print("Specificity")
5401/(5401+389)

# Sensitivity
print("Sensitivity")
1828/(1828+1431)

# Precision
print("Precision")
1828/(1828+389)

# Error rate
print("Error rate")
(389+1431)/(1828+5401+1431+389)

```

```{r}
summary(logic_model$residuals)
```

```{r}
plot(predict(logic_model),residuals(logic_model))
abline(h=0,lty=2,col="grey")
```


# TREE

```{r}
treeFit<- rpart(income~.,data=train,method = 'class')
print(treeFit)

```
```{r}
rpart.plot(treeFit, box.col=c("#DD8D29", "#46ACC8"))
```
```{r}
Prediction1<- predict(treeFit,newdata=test[-15],type = 'class')
confusionMatrix(Prediction1,test$income)
TreeAcu<-confusionMatrix(Prediction1,test$income)$overall[1]
TreeAcu
```

```{r}
recall(factor(Prediction1), factor(test$income))
precision(factor(Prediction1), factor(test$income))
F_meas(factor(Prediction1), factor(test$income))
```

```{r}
# USING GINI INDEX
tc<- trainControl(method="cv", number=5)
model2_res<- train(income ~ .,data = train,method="rpart2",
                 parms = list(split = 'gini'),
                 trControl=tc)

rpart.plot(model2_res$finalModel, box.palette = c("#DD8D29", "#46ACC8"))

prediction2_res<- predict(model2_res, test, type="raw")


confusionMatrix(factor(prediction2_res), factor(test$income))
GiniAcu <- confusionMatrix(factor(prediction2_res), factor(test$income))$overall[1]

recall(factor(prediction2_res), factor(test$income))
precision(factor(prediction2_res), factor(test$income))
F_meas(factor(prediction2_res), factor(test$income))
```

```{r}
# USING NAIVE BAYES
model3_res <- naiveBayes(as.factor(income) ~.,
                       data= train,
                       control= tc)

prediction3_res<- predict(model3_res, test)


confusionMatrix(factor(prediction3_res),factor(test$income))
BayesAcu<-confusionMatrix(factor(prediction3_res),factor(test$income))$overall[1]
BayesAcu
recall(factor(prediction3_res), factor(test$income))
precision(factor(prediction3_res), factor(test$income))
F_meas(factor(prediction3_res), factor(test$income))
```

```{r}
Accuracy <- data.frame(Model=c('Knn', 'Logistic Regression','Decision Tree','Gini index','Naive Bayes'),Accuracy=c(acu.knn20,lgAcu, TreeAcu,GiniAcu,BayesAcu))

acc_models <- ggplot(Accuracy,aes(x=Model,y=Accuracy)) +
  geom_bar(stat = 'identity', aes(fill=Model)) +
  ggtitle('Accuracies of Models')+
  theme(legend.position="none")

acc_models <- acc_models+ scale_fill_manual(values=wes_palette(n=5, name="FantasticFox1"))
acc_models
#ggsave("acc_models.png", plot = acc_models, path = "") #Set path
```

```{r}
# EVALUATE MODELS USING ROC CURVE

tree_1 <- Prediction1
Gini_Index_2 <- prediction2_res
Naive_Bayes_2 <-prediction3_res
```

```{r}
colAUC(X=cbind(tree_1, Gini_Index_2, Naive_Bayes_2),
       y=test$income,plotROC=TRUE)
```

```{r}
colAUC(X=cbind( Naive_Bayes_2),
       y=test$income,plotROC=TRUE)
```

```{r}
Random_Forest <- randomForest(income~.,data = train)
Random_Forest
pred_Random_Forest<- predict(Random_Forest,test)

#confusion matrix
confusionMatrix(pred_Random_Forest, test$income, positive = ">50K")
```
Now accuracy is little bit low (83%) as compared to previous model . But our sensitivity is now high 0.829 ans specificity is 0.83. So if our aim is to predict high income people ,our second model should be choosen as sensitivity is high than previous model...This model is predicting high income more correctly.

```{r}
varImpPlot(Random_Forest)
```

